---
title: "010_Regression"
subtitle: "R Notebook"
author: "DataPhilly Workshop"
date: "`r Sys.time()`"
output: html_notebook
---

# Install packages / Load libraries

```{r pkgs, message = FALSE}
if(!require(pacman)) install.packages("pacman")
pacman::p_load("tidyverse", "magrittr", "ggthemes", "caret",
               "corrplot", "IDPmisc", "randomForest", "e1071", "class", "stats", "utils")
```

# DataPhilly 'branding'

```{r branding, message = FALSE}
DataPhilly_colors <- c('#004785', # Wharton blue
                       '#f74902', # Flyers orange
                       '#a90533', # Wharton red
                       '#004953', # Eagles (midnight) green
                       '#ffc600', # Drexel yellow
                       '#c1c6c8'  # Temple silver
)

DataPhilly_theme <- function(base_family = "serif",
                             base_size = 12,
                             plot_title_family = 'sans',
                             plot_title_size = 20,
                             plot_title_color = DataPhilly_colors[1],
                             grid_col = '#dadada') {
  # use these variables
  aplot <-
    ggplot2::theme_minimal(base_family = base_family, base_size = base_size) #piggyback on theme_minimal
  aplot <- aplot + theme(panel.grid = element_line(color = grid_col))
  aplot <-
    aplot + theme(
      plot.title = element_text(
        size = plot_title_size,
        family = plot_title_family,
        color = plot_title_color
      )
    )
  aplot
}
```

# Read data

```{r read_data}
df <- read.csv('data/BCF.csv', stringsAsFactors = FALSE) %>%
  NaRV.omit()

# bookkeeping
colnames(df)[1] <- 'ID'
colnames(df)[2] <- 'EndPt'
descriptors <- colnames(df)[3:ncol(df)]
```

# Plot distribution of endpoint values

```{r plot_orig_data}
ggplot2::ggplot(df, aes(x = EndPt)) +
  geom_histogram(binwidth = 0.1, fill = DataPhilly_colors[1], colour = "white") +
  scale_fill_manual(values = DataPhilly_colors) +
  DataPhilly_theme() +
  labs(
    title = 'log(BCF) Endpoint Distribution',
    subtitle = 'BCF = Bioconcentration Factor',
    x = 'log(BCF)',
    y = 'Count',
    caption = 'DataPhilly Workshop: StoredDataStories'
  )
```

# Build stratified testing and training datasets

```{r train_test}
inTrain = caret::createDataPartition(df$EndPt, p = 0.8, list = FALSE)
dfTrain <- df[inTrain, ]
dfTest <- df[-inTrain, ]

dfTrain$set <- 'train'
dfTest$set <- 'test'

df <- rbind(dfTrain, dfTest)

ggplot(df, aes(x = EndPt,  colour = set)) +
  geom_density() +
  scale_colour_manual(values = DataPhilly_colors) +
  DataPhilly_theme() +
  labs(
    title = 'log(BCF) Endpoint Distribution: Train & Test Sets',
    subtitle = 'BCF = Bioconcentration Factor',
    x = 'log(BCF)',
    y = "Density",
    caption = 'DataPhilly Workshop: StoredDataStories'
  )

X_train <- dfTrain %>%
  select(-ID, -EndPt, -set)
y_train <- dfTrain %>%
  select(EndPt) %>%
  data.frame()

X_test <- dfTest %>%
  select(-ID, -EndPt, -set)
y_test <- dfTest %>%
  select(EndPt) %>%
  data.frame()
```

# Curate data

## remove near-zero variance descriptors

```{r nzv}
nzv <- caret::nearZeroVar(X_train, freqCut = 100/0)
colnames(X_train[ , nzv])
X_train <- X_train[ , -nzv]
### and
X_test <- X_test[ , -nzv]
```

## remove highly correlated descriptors

```{r highCorr}
correlations <- cor(X_train)
corrplot::corrplot(correlations, order = 'hclust')
highCorr <- findCorrelation(correlations, cutoff = 0.85)
colnames(X_train[ , highCorr])
X_train <- X_train[ , -highCorr]
### and
X_test <- X_test[ , -highCorr]

correlations <- cor(X_train)
corrplot::corrplot(correlations, order = 'hclust')
```

## center and scale descriptors

```{r center_scale}
preProcValues <- preProcess(X_train, method = c("center", "scale"))

X_trainTransformed <- predict(preProcValues, X_train)
### and
X_testTransformed <- predict(preProcValues, X_test)
```

# Build predictive models

```{r X_y_merged}
train <- cbind(y_train, X_trainTransformed)
test <- cbind(y_test, X_testTransformed)
```

## Random Forest Regression

```{r rf}
rf <- randomForest(EndPt ~ ., data = train, importance = TRUE)

rf.pred <- predict(rf, X_testTransformed)

y_predict <- predict(rf, newdata = X_testTransformed) %>%
  data.frame()
colnames(y_predict) <- c('Predicted')

data2plot <- cbind(y_test, y_predict)

summary(lm(Predicted ~ EndPt, data = data2plot))

rf_plot <-
  ggplot(data2plot, aes(EndPt, Predicted)) +
  geom_point(colour = DataPhilly_colors[1], size = 2) +
  geom_smooth(method = 'lm') +
  labs(title = 'Random Forest') +
  scale_colour_manual(values = DataPhilly_colors) +
  DataPhilly_theme()
rf_plot <- rf_plot + geom_abline(intercept = 0,
                                 slope = 1,
                                 colour = 'red')
rf_plot
```

## Support Vector Machine Regression

```{r svm}
sVect <- svm(
  EndPt ~ .,
  data = train,
  method = "C-classification",
  kernal = "radial",
  gamma = 0.1,
  cost = 10
)

y_predict <- predict(sVect, newdata = X_testTransformed) %>%
  data.frame()
colnames(y_predict) <- c('Predicted')

data2plot <- cbind(y_test, y_predict)

summary(lm(Predicted ~ EndPt, data = data2plot))

svm_plot <-
  ggplot(data2plot, aes(EndPt, Predicted)) +
  geom_point(colour = DataPhilly_colors[1], size = 2) +
  geom_smooth(method = 'lm') +
  labs(title = 'SVM') +
  scale_colour_manual(values = DataPhilly_colors) +
  DataPhilly_theme()
svm_plot <- svm_plot + geom_abline(intercept = 0,
                                 slope = 1,
                                 colour = 'red')
svm_plot
```

## k-Nearest Neighbor Regression

```{r kNN}
neighbors <- knn(X_trainTransformed, X_testTransformed, cl = dfTrain$EndPt, k = 5)

y_predict <- neighbors %>%
  as.character() %>%
  as.numeric() %>%
  data.frame()
colnames(y_predict) <- c('Predicted')

data2plot <- cbind(y_test, y_predict)

summary(lm(Predicted ~ EndPt, data = data2plot))

neighbors_plot <-
  ggplot(data2plot, aes(EndPt, Predicted)) +
  geom_point(colour = DataPhilly_colors[1], size = 2) +
  geom_smooth(method = 'lm') +
  labs(title = 'k-Nearest Neighbors',
       subtitle = 'k = 5') +
  scale_colour_manual(values = DataPhilly_colors) +
  DataPhilly_theme()
neighbors_plot <- neighbors_plot + geom_abline(intercept = 0,
                                               slope = 1,
                                               colour = 'red')
neighbors_plot
```

